Unprepared colonoscopy identifies microbiota specific to the proximal and distal human colon

### 2017-10-17 submission/old_sub_docs/test.tex
text edits made, manuscript getting closer

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The microbiota of the proximal and distal human colon},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{The microbiota of the proximal and distal human colon}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\renewcommand{\familydefault}{\sfdefault}
\usepackage{graphicx}

@@

Running title: The microbiome of the proximal and distal human colon

@@

Kaitlin J. Flynn\textsuperscript{1}, Charles C.
Koumpouras\textsuperscript{1}, Mack T. Ruffin IV\textsuperscript{2},
Danielle Kimberly Turgeon\textsuperscript{3\(\dagger\)}, and Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

@@

\(\dagger\) Corresponding authors:
\href{mailto:kturgeon@med.umich.edu}{\nolinkurl{kturgeon@med.umich.edu}}
and \href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

@@

@@

Chemical and nutrient gradients along the human colon create
microenvironments that affect the distrubution and composition of the
gut microbiota. The microbiome has been implicated in colorectal cancer
(CRC) and inflammatory bowel disease (IBD). Further, these diseases
exhibit different symptoms depending on the location of the colon they
are found in. CRC tumors of the proximal and distal colon are
morphologically and genetically distinct. Similarly, inflammatory bowel
diseases such as Crohn's are typically exacerbated in the proximal
instestine while ulcerative colitis patients often experience symptoms
in the distal colon. Previous analysis of the fecal microbiota from
healthy and CRC or IBD patients has revealed different microbial
signatures associated with these diseases. We extended these
observations of the fecal microbiome to include analysis of the proximal
and distal healthy human colon. We used a two-colonoscope approach on
subjects that had not undergone standard bowel preparation procedure. In
contrast to previous efforts using prepared colonoscopy, this technique
allowed us to characterize the native proximal and distal luminal and
mucosal microbiome without prior chemical disruption. 16S rRNA gene
sequencing was performed on proximal and distal mucosal and luminal
biopsies and home-collected stool for 20 healthy individuals. Diversity
analysis revealed that each site contained a diverse community and that
a patient's samples were more similar to each other than to that of
other individuals. A patient's feces were most similar to samples taken
from the distal lumen, likely reflecting the anatomical structure of the
colon. Since we could not differentiate the communities in the proximal
and distal colon based on community structure or community membership
alone, we employed the Random Forest machine-learning algorithm to
identify key taxa that the two sites in the lumen and mucosa. Random
Forest classification models were built using taxa abundance and sample
location and revealed distinct populations that were found in each
location. A model differentiating the proximal mucosa and lumen was
built with an AUC of 0.856. The proximal mucosa had a higher abundance
of the genera \emph{Enterobacteriaceae} and \emph{Bacteriodes}.
\emph{Peptoniphilus, Anaerococcus, Finegoldia,} and \emph{Turicibacter}
were most likely to be found in distal mucosal samples versus distal
luminal samples (AUC = 0.975). The classification model performed well
(AUC = 0.920) when classifying mucosal samples into proximal or distal
sides, but separating luminal samples from each side proved more
challenging (AUC = 0.575). The distal mucosa was found to have high
populations of \emph{Finegoldia, Murdochiella} and \emph{Porphyromonas}.
Proximal and distal luminal samples were comprised of many of the same
taxa, likely reflecting the fact that stool moves along the colon from
the proximal to distal end. By sampling the unprepped human colon, our
results have identified distinct bacterial populations native to the
proximal and distal sides. Further investigation of these bacteria may
elucidate if and how these groups contribute to different pathogenesis
processes on the respective sides of the colon.

\paragraph{wordcount?}\label{wordcount}

@@

The human colon is an ecosystem comprised of several different
microenvironments inhabited by resident bacterial members of the
microbiome. Concentrations of oxygen, water and anti-microbial peptides
change along the gut axis and influence what populations of microbes
reside in each location. Microenvironments differ not only
longitutinally along the colon, but latitudinally from the epithelium to
mucosa to intestinal lumen, offering several sites for different
microbial communities to flourish. The identity of these specific
microbes and communities are important for understanding the etiology of
complex colon diseases such as Inflammatory Bowel Disease (IBD) and
Colorectal Cancer (CRC). IBD and CRC are known to be preceded or
accelerated by perturbations in gut microbes (1--3). The severity,
symptoms, morbidity and mortality of these diseases is known to vary
based upon the biogeographical location in which they occur. For
instance, CRC tumors that arise on the distal side of the colon are
infiltrating lesions that present with painful symptoms (4). In
contrast, 47\% of CRCs are caused by proximal-sided colon tumors that
are sessile and form along the wall of the colon, often remaining
asymptomatic until advanced carcinogenesis (4). The distal and proximal
sides of the colon differ in the amount of inflammation present and the
genomic instability of precancerous cells, respectively, in addition to
variation in the previously mentioned chemical gradients (1, 5, 6). In
IBD patients, disease flares in the distal colon are usually indicative
of ulcerative colitis (UC), whereas Crohn's disease (CD) patients
typically experience disease in the small intestine, ileum and proximal
colon (2). UC presents as large and highly inflammed mucosal ulcers,
where as CD lesions are often smaller and have areas of normal tissue
distributed amongst the flare (2). Thus, given the varied physiology of
the proximal-distal axis of the colon and known differences in disease
patterns at these sites, symbiotic microbes and their metabolites likely
vary as well, and may influence the heterogenous disease prognoses of
IBD and CRC. Because CRC can be a long-term complication of IBD, the
distribution of microbes is important to understanding the
pathophysiology of both diseases.

Several recent findings have shown that development and progression of
IBD or CRC can be attributed to specific molecular events as a result of
interactions between the gut microbiota and human host (1, 3). For
instance, comparison of the bacteria present on CRC tumors with those
found on nearby healthy tissue has identified specific species that are
tumor-associated (7). These species include the oral pathogens
\emph{Fusobacterium nucleatum} and \emph{Porphyromonas asacharolytica}.
Interestingly, these periodontal pathogens have been highly predictive
of whether a patient had CRC tumors or not in our prior human stool
classification studies (8). \emph{F. nucleatum} has also been found to
be elevated in the stool and biopsies of patients with IBD as compared
to healthy controls (9). Furthermore, studies of \emph{F. nucleatum}
isolated from mucosal biopsies showed that more invasive \emph{F.
nucleatum} positively correlates with IBD disease level (9). Like many
intestinal pathogens, the bacteria appear to have a high-impact despite
being lowly-abundant in the community (2). The physiology of these rare
taxa may contribute to the colonic disease state. These studies often
examined only shed human stool or the small intestine, preventing
fine-resolution analysis of paired samples from the proximal and distal
sides of the colon. Similarly, comparisons of on- or off-tumor/lesion
bacteria rarely have matched tissue from the other side of the colon
from the same patient, limiting what conclusions can be drawn about the
colonic microbiome overall, let alone at that specific site. Due to
these limitations, the contribution of the gut microbiota to IBD and CRC
disease location in the colon is largely undefined. Characterizing these
communities could provide needed insight into disease etiology,
including how the disruption of the healthy community could promote the
initiation or proliferation of the distinct proximal and distal CRC
tumors or IBD flares.

The few existing profiles of the microbial biogeography of the gut have
been limited by sample collection methods. The majority of human gut
microbiome studies have been performed on whole shed feces or on samples
collected during colonoscopy procedures. While the latter method allows
investigators to acquire samples from inside the human colon, typically
this procedure is preceded by the use of bowel preparation methods such
as the consumption of laxatives to cleanse the bowel. Bowel preparation
is essential for detecting cancerous or precancerous lesions in the
colon, but complicates microbiome profiling as the chemicals strip the
bowel of contents and disrupt the mucosal layer (10, 11). As such, what
little information we do have about the biogeographical distribution of
the microbes in the proximal and distal colon is confounded by the bowel
preparation procedure.

Here we address the limitations of previous studies and effectively
identify the microbes specific to the lumen and mucosa of the proximal
and distal healthy human colon. We used an unprepared colonoscopy
technique to sample the natural community of each location of the gut
without prior disruption of the native bacteria in 20 healthy
volunteers. To address the inherent inter-individual variation in human
microbiomes, we used a machine-learning classification algorithm trained
on curated 16S rRNA sequencing reads to identify microbes specific to
each location. We found that our classification models were able to
separate mucosal and lumenal samples as well as differentiate between
sides of the colon based on populations of specific microbes. By
identifying the specific microbes we are poised to ask if and how the
presence or disruption of the microbes at each site contribute to the
development of the specific tumor subtypes of CRC in the proximal and
distal human colon.

@@

Lumenal and mucosal samples were collected from the proximal and distal
colon of 20 healthy humans that had not undergone bowel preparation
(Figure 1). Participants also collected stool at home one week prior to
the procedure. To characterize the bacterial communities present at
these sites, 16S rRNA gene sequencing was performed on extracted DNA
from each sample. As expected, each site was primarily dominated by
\emph{Firmicutes} and \emph{Bacteriodetes} (Figure 2A) (12). Samples had
varying levels of diversity at each site, irrespective of the individual
(Figure 2B). For example, the proximal mucosa was more diverse than the
distal for some individuals while the opposite was true for others.
Therefore we could not identify a clear pattern of changes in microbial
diversity along the gut axis.

To compare similarity between sides (proximal or distal) or sites (lumen
or mucosa), we calculated distances from Operational Taxonomic Unit
(OTU) abundances and compared these distances for all individuals.
Again, across all patient samples we observed a range of distances when
comparing sample locations (Figure 3A) and again those ranges did not
follow a clear pattern on an individual basis. However, when comparing
median distances between the proximal lumen and mucosa, the proximal
versus distal lumen, the proximal versus distal mucosa, and the distal
lumen and mucosa, we found that the proximal lumen and mucosa were most
similar to each other than the other samples (P \textless{} 0.005,
Wilcoxon, BH adjustment).

\paragraph{Fecal samples resemble lumenal samples from the distal
colon}\label{fecal-samples-resemble-lumenal-samples-from-the-distal-colon}

Next, we calculated distances to examine how each sample compared to the
home-collected feces. Amidst variability between patients, we did
identify significantly smaller distance between the distal lumenal
sample and the feces (Figure 3B, P \textless{} 0.05, Wilcoxon, BH
adjustment). Furthermore, there was an even larger difference in the
comparisons of the distal mucosa to the feces, indicating that the
mucosa is different from the stool as compared to lumen (P \textless{}
0.0005, Wilcoxon, BH adjustment). To determine what factors may be
driving the differences seen among the samples, we compared distances
between samples from all subjects (interpersonal) versus samples from
within one subject (intrapersonal). We found that samples from one
individual were far more similar to each other than to other study
subjects (Figure 3C), consistent with previous human microbiome studies
that have sampled multiple sites of the human colon (13--15). Thus
interpersonal variation between subjects drove the differences between
samples more than sample site or location. Overall, the results
comparing the structure of the communities suggest that the contents of
the distal lumen are most representative of the patient's feces, and the
microbes remaining on the mucosa are more distinct.

\paragraph{Random Forest classification models identify important
Operational Taxonomic Units (OTUs) on each
side}\label{random-forest-classification-models-identify-important-operational-taxonomic-units-otus-on-each-side}

To identify OTUs that were distinct at each biogeographical site, we
constructed several Random Forest models trained using OTU abundances.
We used 10-fold cross validation to build the first model to classify
the lumenal versus mucosal samples for the proximal and distal sides,
independently (Figure S1A). The models were constructed to only use the
five most predictive OTUs as input to reduce overfitting between
samples. The models performed well when classifying these samples (AUC =
0.856 and AUC = 0.975, respectively). The OTUs that were most predictive
of each site were identified by their greatest mean decrease in accuracy
when removed from the model. For distinguishing the proximal lumen and
mucosa, OTUs from the \emph{Bacteriodes}, \emph{Actinomyces},
\emph{Psuedomonas} and two OTUs from the \emph{Enterobacteraceae} genera
were differentially abundant (Figure 4A). The model classifying the
distal lumen and mucosa identified OTUs from \emph{Turicibacter},
\emph{Finegoldia}, \emph{Peptoniphilus} and two OTUs from the
\emph{Anaerococcus} genera (Figure 4B). These results indicated that
there were fine differences between the different sites of the colon,
and that these could be traced to specific OTUs on each side.

Next, we built a model to differentiate the proximal and distal lumenal
samples using 10-fold cross validation. The model performed best when
distinguishing the proximal versus distal mucosa (Figure S1B, AUC =
0.920) compared to the proximal versus distal lumen (AUC = 0.575). These
models were also constructed to use only the five most predictive OTUs
as input to reduce overfitting. OTUs that were differentially abundant
between the distal and proximal mucosa included members of the
\emph{Porphyromonas}, \emph{Murdochiella}, \emph{Finegoldia},
\emph{Anaerococcus} and \emph{Peptoniphilus} genera (Figure 5A).
Differentially abundant OTUs of the proximal and distal lumen included
three OTUs of the \emph{Bacteroides} genus, a \emph{Clostridium IV} OTU
and an \emph{Oscillibacter} OTU (Figure 5B). This analysis found that
some of the same OTUs that are distinct between the mucosa and lumen
also helped to differentiate between the two sides- such as
\emph{Anaerococcus} and \emph{Finegoldia}.

@@

Given that specific bacterial species have been associated with
colorectal cancer and IBD, we probed our sample set for these OTUs.
Among our 100 samples, the most frequent sequence associated with the
\emph{Fusobacterium} genus was OTU179, which aligns via BLASTn to
\emph{Fusobacterium nucleatum subsp animalis} (100\% over full length).
This is the only species of \emph{Fusobacterium} known to have oncogenic
properties and be found on the surfaces of colorectal cancer tumors.
(16). There were 14 patient samples with the \emph{F. nucleatum subsp.
animalis} sequences. Of the samples with the highest abundance of
\emph{F. nucleatum subsp. animalis}, four of the samples were from the
proximal mucosa and three from the distal mucosa (Figure 6A). The second
most frequent \emph{Fusobacterium} sequence was OTU472, which aligned
with 99\% identity to \emph{F. varium}. In addition to \emph{F.
nucleatum}, \emph{F. varium} has been associated with IBD (17). Four
study participants harbored \emph{F. varium} and the samples were split
evenly between the proximal and distal mucosa (Figure 6B). OTU152 was
similar to the members of the \emph{Porphyromonas} genus and the most
frequent sequence in that OTU aligned to \emph{Porphyromonas
asacharolytica} (99\% over full length), another bacterium commonly
detected and isolated from colorectal tumors. OTU152 was only detected
on the distal mucosa, and in fact was one of the OTUs the classification
model identified as separating distal and proximal sides (Figure 6C).
Among the 11 distal mucosa samples that were positive for \emph{P.
asacharolytica}, the relative abundances for this OTU ranged from 0.01\%
- 16\%. Thus, disease-associated OTUs could be found in our sample set
of 20 healthy individuals.

@@

Here we identified bacterial taxa that were specific to the lumen and
mucosa of the proximal and distal sides of the human colon from samples
collected during unprepared colonoscopy. We found that all locations
contained a range of phyla abundances and a range of diversity, but that
there was a wide variability between subjects. Pairwise comparisons of
each of the sites revealed that the proximal mucosa and lumen were most
similar to each other. Further, comparison of colonoscopy-collected
samples with samples collected from stool at home showed that the distal
lumen was most similar to feces. Random Forest models built on OTU
relative abundances from each sample identified microbes that were
particular to each location of the colon. Finally, we were able to
detect some bacterial OTUs associated with colonic disease in our
healthy patient cohort. Using unprepped colonoscopy and machine
learning, we have identified bacterial taxa specific to the healthy
proximal and distal human colon.

When examining the relative abundance of the dominant phyla at each site
(i.e. \emph{Bacteriodes} and \emph{Firmicutes}), there was a wide amount
of variation. This likely reflects not only the variability between
human subjects, casued by differences in age, sex, diet, but also
biogeogrphical ``patchiness'' in the gut microbiome. Several studies
have noted that the bacteria recoverable from the same mucosal sample
location can be vastly different when the samples are taken just 1 cm
away from each other (18). Similar patchiness is also observed in
lumenal contents and fecal samples themselves; there is observed
separation of different interacting microbes along the length of a stool
sample, for instance (19). That said, across our samples the mucosal
samples harbor more \emph{Proteobacteria}, consistent with previous
studies comparing mucosal swabs to lumenal content in humans (5). Hence,
the conclusions we can draw from phyla analysis may be impacted by
patchiness between subjects.

>>>>

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Unprepared colonoscopy identifies microbiota specific to the proximal and distal human colon},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setlength{\droptitle}{-2em}
  \title{Unprepared colonoscopy identifies microbiota specific to the proximal
and distal human colon}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{}
  \preauthor{}\postauthor{}
  \date{}
  \predate{}\postdate{}

\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers

@@

Running title: Unprepared colonoscopy identifies proximal and distal
human colonic microbiota

@@

Kaitlin J. Flynn\textsuperscript{1}, Charles C.
Koumpouras\textsuperscript{1}, Mack T. Ruffin IV\textsuperscript{2},
Danielle Kimberly Turgeon\textsuperscript{3}, and Patrick D.
Schloss\textsuperscript{1\(\dagger\)}

@@

\(\dagger\) Corresponding author:
\href{mailto:pschloss@umich.edu}{\nolinkurl{pschloss@umich.edu}}

@@

\begin{verbatim}
## Warning in roc.default(response, m[[predictors]], ...): 'response' has
## more than two levels. Consider setting 'levels' explicitly or using
## 'multiclass.roc' instead
\end{verbatim}

@@

The human colon contains chemicals and nutrients that change along the
proximal to distal gut axis. These gradients create microenvironments
that affect the distrubution and composition of the gut microbiota. The
microbiome has been implicated in the colonic diseases colorectal cancer
(CRC) and inflammatory bowel disease (IBD). Further, these diseases
exhibit different symptoms depending on the location of the colon they
are found in. CRC tumors of the proximal and distal colon are
morphologically and genetically distinct. Similarly, bowel diseases such
as Crohn's are typically exacerbated in the proximal instestine while
ulcerative colitis patients often experience symptoms in the distal
colon. Previous analysis of the fecal microbiota from healthy and CRC or
IBD patients has revealed different microbial signatures associated with
these diseases. We extended these observations of the fecal microbiome
to include analysis of the proximal and distal healthy human colon. We
used a two-colonoscope approach on subjects that had not undergone
standard bowel preparation procedure. This technique allowed us to
characterize the native proximal and distal luminal and mucosal
microbiome without prior chemical disruption. 16S rRNA gene sequencing
was performed on proximal and distal mucosal and luminal biopsies and
home-collected stool for 20 healthy individuals. Diversity analysis
revealed that each site contained a diverse community, and that a
patient's samples were more similar to each other than to that of other
individuals. Comparison of all samples to fecal samples taken at exit
uncovered that the feces were most similar to samples taken from the
distal lumen, likely reflecting the anatomical structure of the colon.
Since we could not differentiate sites along the colon based on
community structure or community membership alone, we employed the
Random Forest machine-learning algorithm to identify key species that
distinguish biogeographical sites. Random Forest classification models
were built using taxa abundance and sample location and revealed
distinct populations that were found in each location.
\emph{Peptoniphilus, Anaerococcus, Enterobacteraceae, Pseudomonas} and
\emph{Actinomyces} were most likely to be found in mucosal samples
versus luminal samples (AUC = 0.925). The classification model performed
well (AUC = 0.912) when classifying mucosal samples into proximal or
distal sides, but separating luminal samples from each side proved more
challenging (AUC = 0.755). The distal mucosa was found to have high
populations of \emph{Finegoldia, Murdochiella} and \emph{Porphyromonas}.
Proximal and distal luminal samples were comprised of many of the same
taxa, likely reflecting the fact that stool moves along the colon from
the proximal to distal end. By sampling the unprepped human colon, our
results have identified distinct bacterial populations native to the
proximal and distal sides. Further investigation of these bacteria may
elucidate if and how these groups contribute to different pathogenesis
processes on the respective sides of the colon.

@@

The human colon is an ecosystem comprised of several different
microenvironments inhabited by resident bacterial members of the
microbiome. Concentrations of oxygen, water and anti-microbial peptides
change along the gut axis and influence what populations of microbes
reside in each location. Microenvironments differ not only
longitutinally along the colon, but latitudinally from the epithelium to
mucosa to intestinal lumen, offering several sites for different
microbial communities to flourish.

\emph{2 sentences in here are broken}

The severity, symptoms, morbidity and mortality of these diseases is
known to vary based upon the biogeographical location in which they
occur. For instance, CRC tumors that arise on the distal side of the
colon are infiltrating lesions that present with painful symptoms ((1)).
In contrast, 47\% of CRCs are caused by proximal-sided colon tumors that
are sessile and form along the wall of the colon, often remaining
asymptomatic until advanced carcinogenesis ((1)). The distal and
proximal sides of the colon differ in the amount of inflammation present
and the genomic instability of precancerous cells, respectively, in
addition to variation in the previously mentioned chemical gradients
((2), (3), (4)). In IBD patients, disease flares in the distal colon are
usually indicative of ulcerative colitis (UC), whereas Crohn's disease
(CD) patients typically experience disease in the small intestine, ileum
and proximal colon ((5)). UC presents as large and highly inflammed
mucosal ulcers, where as CD lesions are often smaller and have areas of
normal tissue distributed amongst the flare ((5)). Thus, given the
varied physiology of the proximal-distal axis of the colon and known
differences in disease patterns at these sites, symbiotic microbes and
their metabolites likely vary as well, and may influence the
heterogenous disease prognoses of IBD and CRC. Because CRC is a
long-term complication of IBD, the distribution of microbes is important
to understanding the pathophysiology of both diseases.

Several recent findings have shown that development and progression of
IBD or CRC can be attributed to specific molecular events as a result of
interactions between the gut microbiota and human host ((2)).

\emph{missing last half of second paragraph here, hasnt been fully
tested}

The few existing profiles of the microbial biogeography of the gut have
been limited by sample collection methods. The majority of human gut
microbiome studies have been performed on whole shed stool or on samples
collected during colonoscopy procedures. While the latter method allows
investigators to acquire samples from inside the human colon, typically
this procedure is preceded by the use of bowel preparation methods such
as the consumption of laxatives to cleanse the bowel. Bowel preparation
is essential for detecting cancerous or precancerous lesions in the
colon, but complicates microbiome profiling as the chemicals strip the
bowel of contents and disrupt the mucosal layer ((6), (7)). As such,
what little information we do have about the biogeographical
distribution of the microbes in the proximal and distal colon is
confounded by the bowel preparation procedure.

Here we aimed to address the limitations of previous studies and
effectively identify the microbes specific to the lumen and mucosa of
the proximal and distal healthy human colon. Our design used an
unprepared colonoscopy technique to sample the natural community of each
location of the gut without prior disruption of the native bacteria in
20 healthy volunteers. To address the inherent inter-individual
variation in human microbiomes, we used a machine-learning
classification algorithm trained on curated 16S rRNA sequencing reads to
identify microbes specific to each location. We found that our
classification models were able to separate mucosal and lumenal samples
as well as differentiate between sides of the colon based on populations
of specific microbes. By identifying the specific microbes we are poised
to ask if and how the presence or disruption of the microbes at each
site contribute to the development of the specific tumor subtypes of CRC
in the proximal and distal human colon.

@@

Lumenal and mucosal samples were collected from the proximal and distal
colon of 20 healthy humans that had not undergone bowel preparation
(Figure 1). Participants also collected stool at home one week prior to
the procedure. To characterize the bacterial communities present at
these sites, 16S rRNA gene sequencing was performed on extracted DNA
from each sample. Each site was primarily dominated by \emph{Firmicutes}
and \emph{Bacteriodetes} (Figure 2A), consistent with known variability
in human microbiome research (cite). Likewise, samples had varying
levels of diversity at each site, irrespective of the individual (Figure
2B). For example, the proximal mucosa was more diverse than the distal
for some individuals while the opposite was true for others. Therefore
we could not identify a clear pattern of changes in microbial diversity
along the gut axis.

To compare similarity between sides (proximal or distal) or sites (lumen
or mucosa), we calculated \({\theta}\)YC distances from OTU abundances
and compared these distances for all individuals. Again, across all
patient samples we observed a range of \({\theta}\)YC distances when
comparing sample locations (Figure 3A) and again those ranges did not
follow a clear pattern on an individual basis. However, when comparing
median distances between the proximal lumen and mucosa, the proximal
versus distal lumen, the proximal versus distal mucosa, and the distal
lumen and mucosa, we found that the proximal lumen and mucosa were most
similar to each other than the other samples (P \textless{} 0.005,
Wilcoxon, BH adjustment).

\paragraph{Stool at exit most resembles lumenal samples from the distal
colon}\label{stool-at-exit-most-resembles-lumenal-samples-from-the-distal-colon}

Next, we calculated \({\theta}\)YC distances to examine how each sample
compared to the home-collected exit stool. Amidst variability between
patients, we did identify significantly smaller \({\theta}\)YC distance
between the distal lumenal sample and the exit stool (Figure 3B, P
\textless{} 0.05, Wilcoxon, BH adjustment). Furthermore, there was an
even larger difference in the comparisons of the distal mucosa to the
exit stool, indicating that the mucosa is different from the stool as
compared to lumen (P \textless{} 0.0005, Wilcoxon, BH adjustment). To
determine what factors may be driving the differences seen among the
samples, we compared thetaYC distances between samples from all subjects
(interpersonal) versus samples from within one subject (intrapersonal).
We found that samples from one individual were far more similar to each
other than to other study subjects (Figure 3C), consistent with previous
human microbiome studies that have sampled multiple sites of the human
colon (\textbf{???}, (8), (9)). Thus interpersonal variation between
subjects drives the differences between samples more than sample site or
location. Overall, the results comparing the structure of the
communities suggest that the contents of the distal lumen are most
representative of stool at exit, and the microbes remaining on the
mucosa are much different.

\paragraph{Random Forest classification models identify important OTUs
on each
side}\label{random-forest-classification-models-identify-important-otus-on-each-side}

To identify OTUs that were distinct at each biogeographical site, we
constructed several Random Forest models trained using OTU abundances.
We built the first model to classify the lumen versus mucosal samples
for the proximal and distal sides, independently (Figure 4A). The
constructed model used ((Xopt)) features for the proximal and ((Xopt))
for the distal. The models performed well when classifying these samples
(0.8312912 and 0.9804871, respectively). The OTUs that were most
predictive of each site are identified by their greatest mean decrease
in accuracy when removed from the model. For distinguishing the proximal
lumen and mucosa, OTUs from the \emph{Bacteriodes}, \emph{Actinomyces},
\emph{Psuedomonas} and two OTUs from the \emph{Enterobacteraceae} genera
were differentially abundant (Figure 4B). The model classifying the
distal lumen and mucosa identified OTUs from \emph{Turicibacter},
\emph{Finegoldia}, \emph{Peptoniphilus} and two OTUs from the
\emph{Anaerococcus} genera that could distinguish lumen from mucosal
samples (Figure 4C). These results indicate that there are fine
differences between the different sites of the colon, and that these can
be traced down to specific OTUs on each side.

Next, we built a model to differentiate the proximal and distal lumenal
samples. The model performed best when distinguishing the proximal
versus distal mucosa (Figure 5A, AUC = 0.912) compared to the proximal
versus distal lumen (AUC = 0.755). These models were able to explain
((X\%)) of the variance, respectively. OTUs that were differentially
abundant between the distal and proximal mucosa included members of the
\emph{Porphyromonas}, \emph{Murdochiella}, \emph{Finegoldia},
\emph{Anaerococcus} and \emph{Peptoniphilus} genera (Figure 5B).
Differentially abundant OTUs of the proximal and distal lumen included
three OTUs of the \emph{Bacteroides} genus, a \emph{Clostridium IV} OTU
and an \emph{Oscillibacter} OTU (Figure 5C). This analysis found that
some of the same OTUs that are distinct between the mucosa and lumen
also helped to differentiate between the two sides- such as
\emph{Anaerococcus} and \emph{Finegoldia}.

@@

Given that specific bacterial species have been associated with
colorectal cancer and IBD, we probed our sample set for these OTUs.
Among our 100 samples, the most frequent sequence associated with the
\emph{Fusobacterium} genus was OTU179, which aligns via BLASTn to
\emph{Fusobacterium nucleatum subsp animalis} (XX\% over full length).
This is the only species of \emph{Fusobacterium} known to have oncogenic
properties and be found on the surfaces of colorectal cancer tumors.
((10)). The \emph{Fusobacterium} positive samples were located in x\% of
the the proximal and X\% of the distal mucosa and represented as much as
1\% of any sample (Figure 6A). OTU152 was similar to the members of the
\emph{Porphyromonas} genus and the most frequent sequence in that OTU
aligned to \emph{Porphyromonas asacharolytica} (X\% over full length),
another bacterium commonly detected and isolated from colorectal tumors.
OTU152 was only detected on the distal mucosa, and in fact was one of
the OTUs the classification model identified as separating distal and
proximal sides (Figure 6B). Among the samples that were positive for
\emph{P. asacharolytica}, the relative abundances for this OTU ranged
from 0.01\% - 16\%. Thus, disease-associated OTUs could be found in our
sample set of 20 healthy individuals.

@@

Here we identified bacterial taxa that were specific to the lumen and
mucosa of the proximal and distal sides of the human colon from samples
collected during unprepared colonoscopy. We found that all locations
contained a range of phyla and a range of diversity, but that there was
a wide variability between subjects. Pairwise comparisons of each of the
sites revealed that the proximal mucosa and lumen were most similar to
each other. Further, comparison of colonoscopy-collected samples with
samples collected from stool at home showed that the distal lumen is
most similar to stool at exit. Random Forest models built on OTU
relative abundances from each sample identified microbes that are
particular to each location of the colon. Finally, we were able to
detect some bacterial OTUs associated with colonic disease in our
healthy patient cohort. Using unprepped colonoscopy and machine
learning, we have identified bacterial phyla specific to the healthy
proximal and distal human colon.

When examining the relative abundance of the different phyla at each
site, there was a wide amount of variation for each phyla with
communities primarily dominated by the \emph{Bacteriodes} and
\emph{Firmicutes}. This likely reflects not only the variability between
human subjects, casued by differences in age, gender, diet, but also
reports of biogeogrphical ``patchiness'' in the gut microbiome. Several
studies have noted that the bacteria recoverable from the same mucosal
sample location can be vastly different when the samples are taken just
1 cm away from each other (11). Similar patchiness is also observed in
lumenal contents and fecal samples themselves; there is observed
separation of different interacting microbes along the length of a stool
sample, for instance (12). That said, across our samples the mucosal
samples harbor more \emph{Proteobacteria}, consistent with previous
studies comparing mucosal swabs to lumenal content in humans (3). Hence,
the conclusions we can draw from phyla analysis are likely impacted by
patchiness between subjects.
